\LARGE \bf Planning with Attitude

### 2021-01-05 paper/main.tex
removed unnecessary frame comment
Update formatting

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper
\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command
\overrideIEEEmargins                                      % Needed to meet printer requirements.
\usepackage[top=60pt, left=48pt, right=48pt, bottom=43pt]{geometry}

@@

\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}
\newcommand{\added}[1]{\textcolor{blue}{#1}}

@@

\author{Brian E. Jackson$^1$, Kevin Tracy$^1$, and Zachary Manchester$^1$%
    \thanks{
        $^1$Robotics Institute, 
        Carnegie Mellon University, 
        5000 Forbes Ave, Pittsburgh, PA, USA
    }
}

@@

@@

    Many robotic systems---including quadrotors, airplanes, satellites, autonomous
    underwater vehicles, and quadrupeds---can perform arbitrarily large three-dimensional
    translations and rotations as part of their normal operation. While representing
    translations is straightforward and intuitive, effectively representing the
    nontrivial group structure of 3D rotations has been a topic of study for many
    decades. Although we can intuitively deduce that rotations are three-dimensional, a
    globally non-singular three-parameter representation of the space of rotations does
    not exist \cite{stuelpnagel1964parametrization}. As a result, when parameterizing
    rotations, we must either a) choose a three-parameter representation and deal with
    singularities and discontinuities, or b) choose a higher-dimensional representation and deal with
    constraints between the parameters. While simply representing attitude is nontrivial,
    generating and tracking motion plans for floating-base systems is an even more
    challenging problem.

@@

    \subsection{Rigid Body Dynamics} \label{sec:rigidbody_dynamics}
        For clarity, we will restrict our attention to rigid bodies moving freely in 3D 
        space. That is, we consider systems with dynamics of the following form:
        \begin{equation} \label{eq:rigid_body_dynamics}
            x = \begin{bmatrix} r \\ q \\ v \\ \omega \end{bmatrix}, \quad 
            \dot{x} = \begin{bmatrix} 
                v \\ 
                \half \q \otimes \hat{\omega} = \half L(q) H \omega \\ 
                \frac{1}{m} \inframe{W}{F(x,u)} \\ 
                J^{-1}(\inframe{B}{\tau(x,u)} - \omega \times J \omega) 
            \end{bmatrix}
        \end{equation}
        where $x$ and $u$ are the state and control vectors, $r \in \R^3$ is the
        position, $\q \in \Q$ is the attitude, $v \in \R^3$ is the linear velocity, and
        $\omega \in \R^3$ is the angular velocity. $m \in \R$ is the mass, $J \in
        \R^{3\times3}$ is the inertia matrix, $\inframe{W}{F(x,u)} \in \R^3$ are the
        forces in the world frame, and $\inframe{B}{\tau(x,u)}$ are the moments in the
        body frame. \added{We use the convention that ``attitude'' always refers to a rotation from body to world frame.}

>>>>

% \documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper
\documentclass[letterpaper, 10 pt, journal, twoside]{ieeetran}
% \IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command
% \overrideIEEEmargins                                      % Needed to meet printer requirements.
% \usepackage[top=60pt, left=48pt, right=48pt, bottom=43pt]{geometry}

@@

\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}
\newcommand{\added}[1]{\textcolor{black}{#1}}

% Header, Title, etc.
\markboth{IEEE Robotics and Automation Letters. Preprint Version. Accepted Dec, 2020}
{Jackson \MakeLowercase{et al.}: Planning with Attitude}

@@

% Make room for more info lines in the \author command 
\author{Brian E Jackson$^{1}$, Kevin Tracy$^{1}$, and Zachary Manchester$^{1}$%
\thanks{Manuscript received: October, 15, 2020; Revised January 2, 2021; Accepted January 4, 2021.}%Use only for final RAL version
\thanks{This paper was recommended for publication by Editor Nancy Amato upon evaluation of the Associate Editor and Reviewers' comments.
This work was supported by a NASA Early Career Faculty Award, NASA JPL, and NSF GFRP Grant No. DGE-1656518.} %Use only for final RAL version
\thanks{$^{1}$All authors are with the Robotics Institute, Carnegie Mellon University, Pittsburgh, PA, USA 
        {\tt\footnotesize brian.jackson@cmu.edu}}%
\thanks{Digital Object Identifier (DOI): see top of this page.}
}

@@

\begin{IEEEkeywords}
    Motion and Path Planning, Optimization and Optimal Control, Computational Geometry, Underactuated Robots, Motion Control, 
\end{IEEEkeywords}

@@

    \IEEEPARstart{M}{any} robotic systems---including quadrotors, airplanes, satellites, autonomous
    underwater vehicles, and quadrupeds---can perform arbitrarily large three-dimensional
    translations and rotations as part of their normal operation. While representing
    translations is straightforward and intuitive, effectively representing the
    nontrivial group structure of 3D rotations has been a topic of study for many
    decades. Although we can intuitively deduce that rotations are three-dimensional, a
    globally non-singular three-parameter representation of the space of rotations does
    not exist \cite{stuelpnagel1964parametrization}. As a result, when parameterizing
    rotations, we must either a) choose a three-parameter representation and deal with
    singularities and discontinuities, or b) choose a higher-dimensional representation and deal with
    constraints between the parameters. While simply representing attitude is nontrivial,
    generating and tracking motion plans for floating-base systems is an even more
    challenging problem.

@@

    \subsection{Rigid Body Dynamics} \label{sec:rigidbody_dynamics}
        For clarity, we will restrict our attention to rigid bodies moving freely in 3D 
        space. That is, we consider systems with dynamics of the following form:
        \begin{equation} \label{eq:rigid_body_dynamics}
            x = \begin{bmatrix} r \\ q \\ v \\ \omega \end{bmatrix}, \quad 
            \dot{x} = \begin{bmatrix} 
                v \\ 
                \half \q \otimes \hat{\omega} = \half L(q) H \omega \\ 
                \frac{1}{m} \inframe{W}{F(x,u)} \\ 
                J^{-1}(\inframe{B}{\tau(x,u)} - \omega \times J \omega) 
            \end{bmatrix}
        \end{equation}
        where $x$ and $u$ are the state and control vectors, $r \in \R^3$ is the
        position, $\q \in \Q$ is the attitude, $v \in \R^3$ is the linear velocity, and
        $\omega \in \R^3$ is the angular velocity. $m \in \R$ is the mass, $J \in
        \R^{3\times3}$ is the inertia matrix, $\inframe{W}{F(x,u)} \in \R^3$ are the
        forces in the world frame, and $\inframe{B}{\tau(x,u)}$ are the moments in the
        body frame.
